---
title: The Effect of the Method of Cutting on the Moisture Content of Samples from
  Tree Branches
geometry: margin=2.5cm
output:
  html_document:
    df_print: paged
    toc: no
  pdf_document:
    highlight: zenburn
    toc: no
fontsize: 12pt
---


## Introduction

Your report should contain a short background/intro to the problem (including the aim of the original study).

A presentation of the results of your 
- statistical analyses
- including exploratory data analysis
- model fitting and final model
- along with a short discussion of any shortcomings of the final model
- your conclusions


## Result

```{r, echo=FALSE}
## this is how you can 
library(reticulate)
file <- read.table('tree-moisture.txt', 
                 col.names = c("Species", "Branch", "Location", "TranspirationType", "Moisture"))
file$Species[file$Species==1]<-"LP"
file$Species[file$Species==2]<-"SP"
file$Species[file$Species==3]<-"YP"
file$Species[file$Species==4]<-"RG"

file$Location[file$Location ==1]<-"Central"
file$Location[file$Location ==2]<-"Distal"
file$Location[file$Location ==3]<-"Proximal"

file$TranspirationType[file$TranspirationType==1]<-"Rapid"
file$TranspirationType[file$TranspirationType==2]<-"Slow"

moisture <-file[,c("Species", "Location", "TranspirationType", "Moisture")]

species <- factor(c(moisture$Species))
location <- factor(c(moisture$Location))
transpiration <- factor(c(moisture$TranspirationType))

moisture$Species <- species
moisture$Location <- location
moisture$TranspirationType <- transpiration

```
### Hypotheses:
- H0: The means of all species groups are equal.
- H1: The mean of at least one species group is different.

- H0: The means of the location groups are equal.
- H1: The mean of at least one species group is different.

- H0: The means of the transpiration type groups are equal.
- H1: The means of the transpiration type groups are different.

- H0: There is no interaction between species and location of cut.
- H1: There is interaction between species and location of cut.

- H0: There is no interaction between species and transpiration type.
- H1: There is interaction between species and transpiration type.

- H0: There is no interaction between location of cut and transpiration type.
- H1: There is interaction between location of cut and transpiration type.

### Exploratory data analysis
#### Means for each combination of **species**, **location**, **transpiration type**.
```{r, echo=FALSE}
file
tapply(moisture$Moisture, list(moisture$Species, moisture$Location, moisture$TranspirationType),mean)
```
#### Standard Deviations for every combination of **species**, **location**, **transpiration type**
```{r}
tapply(moisture$Moisture, list(moisture$Species, moisture$Location, moisture$TranspirationType), sd)
```
```{r}
tapply(moisture$Moisture, list(moisture$Species, moisture$Location, moisture$TranspirationType), median)
```
```{r}
tapply(moisture$Moisture, list(moisture$Species, moisture$Location, moisture$TranspirationType), min)
```
```{r}
tapply(moisture$Moisture, list(moisture$Species, moisture$Location, moisture$TranspirationType), max)
```
```{r}
utils::str(moisture)
plot.design(moisture)
```

##### ANOVA Models
```{r}
m_specieslocation.aov <- aov(Moisture ~ Species*Location, data=moisture) 
summary(m_specieslocation.aov)

m_speciestranspiration.aov <- aov(Moisture ~ Species*TranspirationType, data=moisture) 
summary(m_speciestranspiration.aov)

m_locationtranspiration.aov <- aov(Moisture ~ Location*TranspirationType, data=moisture) 
summary(m_locationtranspiration.aov)
```
##### coefficients

```{r}
coef(m_specieslocation.aov)
coef(m_speciestranspiration.aov)
coef(m_locationtranspiration.aov)

```
##### INTERACTION PLOTS

```{r}
interaction.plot(moisture$Location, moisture$Species, moisture$Moisture)
interaction.plot(moisture$TranspirationType, moisture$Species, moisture$Moisture)
interaction.plot(moisture$TranspirationType, moisture$Location, moisture$Moisture)

```
```{r}
interaction.plot(moisture$Species, moisture$Location, moisture$Moisture)
interaction.plot(moisture$Species, moisture$TranspirationType, moisture$Moisture)
interaction.plot(moisture$Location, moisture$TranspirationType, moisture$Moisture)

```

```{r}
layout(matrix(1:4,ncol=2)) 

plot(m_specieslocation.aov)
plot(m_speciestranspiration.aov)
plot(m_locationtranspiration.aov)
```

### 2-way ANOVA
#### 3 2-way ANOVA tests
##### Species vs. Location


##### Species vs Transpiration Type



##### Location vs Transpiration Type
## Summary

